paper figures

### 2025-01-31 paper_figures.Rmd
some additional plots

```{r Fig4D, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale} 
FB.data.pooled <- read.csv('data/FB_data_pooled.csv', header=1, row.names=1) %>%
  dplyr::mutate(num_variants_somatic = num_variants * (1 - frac_variants_germline),
                status=ifelse(Diagnosis %in% c('Control','Subclass','Atypical'), 'Control', 'CVID')) %>%
  tibble::column_to_rownames('Sample')
QP.data <- read.csv('data/QP_data.csv', header=1, row.names=1) %>%
  dplyr::mutate(Diagnosis=plyr::revalue(Diagnosis,c('control'='Control'))) %>%
  tibble::column_to_rownames('Sample')

FB_train_donors <- paste0('FB_',sprintf('%0.2d', 1:35))
human.meta <- rbind(FB.data.pooled %>%
                      dplyr::mutate(dataset=ifelse(grepl('FB',Donor),
                                                   ifelse(Sample_donor %in% FB_train_donors,'training','testing'),
                                                   'validation')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint, dataset,Sex,Age,complication, EUROclass, pct_B, pct_IgA, pct_IgG),
                    QP.data %>%
                      dplyr::mutate(dataset='validation',
                                    Timepoint='t1',
                                    complication=NA,
                                    EUROclass=NA,
                                    pct_B=NA,
                                    pct_IgA=NA,
                                    pct_IgG=NA,
                                    status=ifelse(Diagnosis=='Control','Control','CVID')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint,dataset,Sex,Age,complication, EUROclass, pct_B, pct_IgA, pct_IgG)) %>%
  dplyr::mutate(Diagnosis=factor(Diagnosis,levels=c('Control','Subclass','Atypical','CVID','ICOS','Kabuki','IgG4def',
                                                    'AID','TNFRSF13B',
                                                    'NFKB2','PI3K','ATM','BRCA1','LIG4','NIPBL')),
                status=factor(status, levels=c('Control','CVID')))

@@

another PCA with only subclass features

```{r FigS4C, fig.width=2.75 * fig_scale, fig.height=1.75 * fig_scale}
FB.cols.subclass <- grep('frac_clusters_S[MGA]+',colnames(FB.data.pooled),value=TRUE)
FB.X.subclass <- FB.data.pooled[FB.data.pooled$Sample_donor %in% FB_train_donors,FB.cols.subclass] %>%
  dplyr::mutate_if(is.numeric,scale)

pca.subclass <- prcomp(FB.X.subclass)
pcadat.subclass <- cbind(pca.subclass$x,human.meta[human.meta$dataset=='training',])

ggplot(pcadat.subclass,aes(x=PC1,y=PC2)) +
  geom_point(aes(color=Diagnosis),size=1) +
  geom_line(aes(group=Donor, color=Diagnosis)) +
  geom_polygon(data = pcadat.subclass %>%
                 dplyr::mutate(status=ifelse(Diagnosis=='Subclass', 'Subclass', as.character(status))) %>%
                 dplyr::group_by(status) %>%
                 dplyr::slice(chull(PC1, PC2)), alpha = 0.25, aes(fill=status)) +
  my_theme() +
  scale_fill_manual(values=diagnosis_colors, guide='none') +
  scale_color_manual(values=diagnosis_colors) +
  guides(color=guide_legend(override.aes = list(shape=15,size=2), ncol=1)) +
  theme(aspect.ratio=pca.subclass$sdev[2]**2/pca.subclass$sdev[1]**2,
        legend.position='right',
        legend.key.size=unit(1.5,'mm'),
        legend.key.spacing.y=unit(0,'mm'),
        legend.box='horizontal') +
  labs(x=paste0('PC 1 (',100*signif(pca.subclass$sdev[1]**2/sum(pca.subclass$sdev**2),3),'%)'),
       y=paste0('PC 2 (',100*signif(pca.subclass$sdev[2]**2/sum(pca.subclass$sdev**2),3),'%)'),
       color='',shape='',fill='')
```

(this PCA uses `r dim(FB.X.subclass)[2]` features)

@@

```{r FigS4B, fig.width=2 * fig_scale, fig.height=1.7 * fig_scale}
subclass.cols <- c("frac_clusters_SA1","frac_clusters_SA2",
                   "frac_clusters_SG1","frac_clusters_SG2","frac_clusters_SG3","frac_clusters_SG4")

@@

```{r FigS4H, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale}
tmp <- FB.data.pooled %>%
  dplyr::mutate(alpha_ratio_FACS=pct_IgA/(pct_IgA + pct_IgG)) %>%
  dplyr::group_by(status) %>%
  dplyr::mutate(status_R=paste0(status," (R=",signif(cor(alpha_ratio_FACS,alpha_ratio_clusters,use='na.or.complete'),2),")"))
ggplot(tmp, aes(x=alpha_ratio_FACS, y=alpha_ratio_clusters, color=status)) + 
  geom_abline(slope=1, intercept=0, size=.5) +
  #stat_cor(method = "pearson", aes(label=..r.label..), size=2.5, show.legend=FALSE, label.x=.75, label.y=c(.4, .33)) +
  geom_point() + 
  my_theme() +
  scale_color_manual(values=diagnosis_colors, breaks=unique(tmp$status), labels=unique(tmp$status_R)) +
  labs(x='FACS', y='SWIBRID', title='IgA / (IgA + IgG)') +
  theme(legend.position=c(.8,.15),
        legend.box=element_blank())
```

@@

@@

@@

```{r FigS4D, fig.width=3 * fig_scale, fig.height=4 * fig_scale}
tmp <- data.frame(pca$rotation) %>%
  tibble::rownames_to_column('col') %>%
  dplyr::left_join(columns, by='col') %>%
  dplyr::mutate(col=factor(col, levels=columns %>% 
                             dplyr::filter(col %in% selected.cols) %>% 
                             dplyr::pull(col)),
                col_name=gsub("_downsampled|_switch",'',col))

@@

```{r Fig4J, fig.width=5.25 * fig_scale, fig.height=1.7 * fig_scale}
features.here <- c('frac_clusters_SG1',
                   'mean_cluster_size_downsampled',
                   'n_homology_switch',
                   'n_untemplated_switch',
                   'frac_breaks_intra',
                   'frac_breaks_duplications')

@@

ggplot(tmp, aes(x=ifelse(grepl('Cluster',group),as.numeric(group),1+as.numeric(group)),
                y=value,fill=group)) +
  geom_boxplot(outlier.shape=NA, lwd=.5) +
  geom_jitter(width=.1,height=0,size=.5) +
  #annotate('text',x=2,y=Inf,label='', size=2) +
  #annotate('text',x=6,y=Inf,label='validation', size=2) +
  # stat_compare_means(comparisons=list(c('Cluster1','Cluster3'),
  #                                     c('Cluster1','Cluster2'),
  #                                     c('Cluster2','Cluster3')),
  #                    method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  # stat_compare_means(comparisons=list(c('Control','ATM'),
  #                                     c('Control','BRCA1')),
  #                    method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(1,2),c(1,3),c(2,3)),
                     method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(5,6),c(5,7),c(6,7)),
                     method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  facet_wrap(.~col,ncol=length(features.here),scales='free_y') +
  my_theme() +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=c(diagnosis_colors,class_colors)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') +
  scale_x_continuous(breaks=c(1,2,3,5,6,7),
                     labels=c('Cluster1','Cluster2','Cluster3','Control','ATM','BRCA1')) +
  labs(x='',y='', color='Diagnosis')
```

@@

```{r FigS4E, fig.width=3 * fig_scale, fig.height=2.5 * fig_scale}
pvals <- list()
for (x in c("Tetanus",
            "Diphteria",
            "PnPS")) {
  for (f in selected.cols) {
    data <- FB.data.pooled[FB.samples,c(x,f)] %>%
      tidyr::drop_na()
    form <- as.formula(paste0(x,'=="negative" ~ ',f))
    res <- glm(form, family='binomial', data=data) %>% summary()
    pvals[[paste0(x,'_',f)]] <- data.frame(pval=coef(res)[2,4],
                                           vacc=x,
                                           col=f)
  }
}

@@

>>>>

```{r Fig4D, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale} 
FB.data.pooled <- read.csv('data/FB_data_pooled.csv', header=1, row.names=1) %>%
  dplyr::mutate(num_variants_somatic = num_variants * (1 - frac_variants_germline),
                status=ifelse(Diagnosis %in% c('Control','Subclass','Atypical'), Diagnosis, 'CVID')) %>%
  tibble::column_to_rownames('Sample')
QP.data <- read.csv('data/QP_data.csv', header=1, row.names=1) %>%
  dplyr::mutate(Diagnosis=plyr::revalue(Diagnosis,c('control'='Control'))) %>%
  tibble::column_to_rownames('Sample')

FB_train_donors <- paste0('FB_',sprintf('%0.2d', 1:35))
human.meta <- rbind(FB.data.pooled %>%
                      dplyr::mutate(dataset=ifelse(grepl('FB',Donor),
                                                   ifelse(Sample_donor %in% FB_train_donors,'training','testing'),
                                                   'validation')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint, dataset,Sex,Age,complication, EUROclass, pct_B, pct_IgA, pct_IgG),
                    QP.data %>%
                      dplyr::mutate(dataset='validation',
                                    Timepoint='t1',
                                    complication=NA,
                                    EUROclass=NA,
                                    pct_B=NA,
                                    pct_IgA=NA,
                                    pct_IgG=NA,
                                    status=ifelse(Diagnosis=='Control','Control','CVID')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint,dataset,Sex,Age,complication, EUROclass, pct_B, pct_IgA, pct_IgG)) %>%
  dplyr::mutate(Diagnosis=factor(Diagnosis,levels=c('Control','Subclass','Atypical','CVID','ICOS','Kabuki','IgG4def',
                                                    'AID','TNFRSF13B',
                                                    'NFKB2','PI3K','ATM','BRCA1','LIG4','NIPBL')),
                status=factor(status, levels=c('Control','Subclass','Atypical','CVID')))

@@

<!-- another PCA with only subclass features -->

<!-- ```{r FigS4C, fig.width=2.75 * fig_scale, fig.height=1.75 * fig_scale} -->
<!-- FB.cols.subclass <- grep('frac_clusters_S[MGA]+',colnames(FB.data.pooled),value=TRUE) -->

<!-- human.X.subclass <- cbind(rbind(FB.data.pooled[,FB.cols.subclass], -->
<!--                                 QP.data[,FB.cols.subclass]), -->
<!--                           human.meta[,'dataset',drop=FALSE]) %>% -->
<!--   tibble::rownames_to_column('sample') %>% -->
<!--   dplyr::mutate(sample=gsub('[0-9]*_|pooled_','',sample)) %>% -->
<!--   dplyr::group_by(dataset) %>% -->
<!--   dplyr::mutate_if(is.numeric,scale) %>% -->
<!--   dplyr::ungroup() %>% -->
<!--   dplyr::select(c('sample',FB.cols.subclass)) %>% -->
<!--   tibble::column_to_rownames('sample') -->

<!-- pca.subclass <- prcomp(human.X.subclass[train,]) -->
<!-- pcadat.subclass <- cbind(pca.subclass$x,human.meta[human.meta$dataset=='training',]) -->

<!-- ggplot(pcadat.subclass,aes(x=PC1,y=PC2)) + -->
<!--   geom_point(aes(color=Diagnosis),size=1) + -->
<!--   geom_line(aes(group=Donor, color=Diagnosis)) + -->
<!--   geom_polygon(data = pcadat.subclass %>% -->
<!--                  #dplyr::mutate(status=ifelse(Diagnosis %in% c('Subclass','Atypical'), as.character(Diagnosis), as.character(status))) %>% -->
<!--                  dplyr::group_by(status) %>% -->
<!--                  dplyr::slice(chull(PC1, PC2)), alpha = 0.25, aes(fill=status)) + -->
<!--   my_theme() + -->
<!--   scale_fill_manual(values=diagnosis_colors, guide='none') + -->
<!--   scale_color_manual(values=diagnosis_colors) + -->
<!--   guides(color=guide_legend(override.aes = list(shape=15,size=2), ncol=1)) + -->
<!--   theme(aspect.ratio=pca.subclass$sdev[2]**2/pca.subclass$sdev[1]**2, -->
<!--         legend.position='right', -->
<!--         legend.key.size=unit(1.5,'mm'), -->
<!--         legend.key.spacing.y=unit(0,'mm'), -->
<!--         legend.box='horizontal') + -->
<!--   labs(x=paste0('PC 1 (',100*signif(pca.subclass$sdev[1]**2/sum(pca.subclass$sdev**2),3),'%)'), -->
<!--        y=paste0('PC 2 (',100*signif(pca.subclass$sdev[2]**2/sum(pca.subclass$sdev**2),3),'%)'), -->
<!--        color='',shape='',fill='') -->
<!-- ``` -->

<!-- (this PCA uses ` dim(human.X.subclass)[2]` features) -->

@@

```{r FigS4D, fig.width=2 * fig_scale, fig.height=1.7 * fig_scale}
subclass.cols <- c("frac_clusters_SA1","frac_clusters_SA2",
                   "frac_clusters_SG1","frac_clusters_SG2","frac_clusters_SG3","frac_clusters_SG4")

@@

```{r FigS4E, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale}
tmp <- FB.data.pooled %>%
  dplyr::mutate(alpha_ratio_FACS=pct_IgA/(pct_IgA + pct_IgG)) %>%
  dplyr::filter(Diagnosis %in% c('Control','CVID','CVID-like')) %>%
  dplyr::group_by(status) %>%
  dplyr::mutate(status_R=paste0(status," (R=",signif(cor(alpha_ratio_FACS,alpha_ratio_clusters,use='na.or.complete'),2),")"))
ggplot(tmp, aes(x=alpha_ratio_FACS, y=alpha_ratio_clusters, color=status)) + 
  geom_abline(slope=1, intercept=0, size=.5) +
  #stat_cor(method = "pearson", aes(label=..r.label..), size=2.5, show.legend=FALSE, label.x=.75, label.y=c(.4, .33)) +
  geom_point() + 
  my_theme() +
  scale_color_manual(values=diagnosis_colors, breaks=unique(tmp$status), labels=unique(tmp$status_R)) +
  labs(x='FACS', y='SWIBRID', title='IgA / (IgA + IgG)') +
  theme(legend.position=c(.8,.15),
        legend.box=element_blank(),
        legend.title=element_blank())
```

@@

<!-- project into PCA using only subclass features -->

<!-- ```{r FigS4X5, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale} -->
<!-- pcadat.combined.subclass <- rbind(cbind(as.matrix(human.X.subclass[train,]) %*% pca.subclass$rotation, human.meta[train,]), -->
<!--                                   cbind(as.matrix(human.X.subclass[test,]) %*% pca.subclass$rotation, human.meta[test,]), -->
<!--                                   cbind(as.matrix(human.X.subclass[validation,]) %*% pca.subclass$rotation, human.meta[validation,])) %>% -->
<!--   dplyr::mutate(dataset=factor(dataset, levels=c('training','testing','validation'))) -->

<!-- ggplot(pcadat.combined.subclass, -->
<!--        aes(x=PC1,y=PC2)) + -->
<!--   geom_polygon(data = pcadat.combined.subclass %>% dplyr::filter(dataset=='training') %>% -->
<!--                  dplyr::group_by(status) %>% -->
<!--                  dplyr::slice(chull(PC1, PC2)), alpha = 0.1, color=NA, aes(fill=status), show.legend=FALSE) + -->
<!--   geom_point(aes(color=Diagnosis, alpha=dataset, shape=dataset), size=1) + -->
<!--   geom_line(aes(group=Donor, color=Diagnosis,alpha=dataset, shape=dataset)) + -->
<!--   my_theme() + -->
<!--   scale_fill_manual(values=diagnosis_colors, guide='none') + -->
<!--   scale_color_manual(values=diagnosis_colors) + -->
<!--   scale_alpha_manual(values=c('testing'=1,'training'=.25,'validation'=1)) + -->
<!--   scale_shape_manual(values=c('testing'=15,'training'=16,'validation'=4)) + -->
<!--   guides(color=guide_legend(override.aes = list(shape=15,size=2), ncol=3), -->
<!--          shape=guide_legend(ncol=1), -->
<!--          alpha=guide_legend(ncol=1)) + -->
<!--   theme(legend.key.height=unit(1.5,'mm'), -->
<!--         legend.key.spacing=unit(0,'mm'), -->
<!--         aspect.ratio=pca$sdev[2]**2/pca$sdev[1]**2, -->
<!--         legend.position='bottom', -->
<!--         legend.box='horizontal') + -->
<!--   coord_cartesian(xlim=c(-3,4), -->
<!--                   ylim=c(-3,2), -->
<!--                   clip='off') + -->
<!--   labs(x='PC 1',y='PC 2', -->
<!--        color='',shape='',fill='',alpha='') -->
<!-- ``` -->

@@

<!-- regression using only subclass features -->

<!-- ```{r FigS4X6, fig.width=2. * fig_scale, fig.height=1.5 * fig_scale, include=FALSE} -->
<!-- set.seed(1) -->
<!-- train.here <- train & !(human.meta[,'Diagnosis'] %in% c('Subclass','Atypical'))  -->
<!-- test.here <- test & !(human.meta[, 'Donor'] %in% FB_train_donors) -->
<!-- cv_model <- cv.glmnet(as.matrix(human.X.subclass[train.here,]), human.meta[train.here,'status']!='Control', alpha = 0, family='binomial') -->
<!-- mod <- glmnet(as.matrix(human.X.subclass[train.here,]), human.meta[train.here,'status']!='Control', alpha = 0, family = "binomial", -->
<!--                 lambda = cv_model$lambda.min) -->
<!-- pred_1 <- predict(mod, type = "response", newx=as.matrix(human.X.subclass[test.here,])) -->
<!-- ROC_1 <- roc(human.meta[test.here,'status']!='Control', pred_1) -->

<!-- pred_2 <- predict(mod, type = "response", newx=as.matrix(human.X.subclass[validation,])) -->
<!-- ROC_2 <- roc(human.meta[validation,'status']!='Control', pred_2) -->

<!-- FB_res <- data.frame(specificity=c(ROC_1$specificities,ROC_2$specificities), -->
<!--                      sensitivity=c(ROC_1$sensitivities,ROC_2$sensitivities), -->
<!--                      thresholds=c(ROC_1$thresholds,ROC_2$thresholds), -->
<!--                      model=c(rep(paste0('training -> testing; AUC=',signif(ROC_1$auc,3)),length(ROC_1$specificities)), -->
<!--                              rep(paste0('training -> validation; AUC=',signif(ROC_2$auc,3)),length(ROC_2$specificities)))) %>% -->
<!--   dplyr::mutate(model=factor(model, levels=unique(model))) %>% -->
<!--   dplyr::arrange(model,sensitivity)  -->

<!-- ggplot(FB_res, aes(x=specificity,y=sensitivity,color=model)) +  -->
<!--   geom_step() + -->
<!--   scale_x_reverse() +   -->
<!--   geom_abline(slope=1,intercept=1,linewidth=.25,linetype='dashed') + -->
<!--   my_theme() + -->
<!--   guides(color=guide_legend(ncol=1)) + -->
<!--   scale_color_manual(values=c('steelblue4','lightsteelblue')) + -->
<!--   theme(legend.position=c(.5,.2)) + -->
<!--   labs(color='',linetype='') -->
<!-- ``` -->

@@

```{r FigS4C, fig.width=3 * fig_scale, fig.height=4 * fig_scale}
tmp <- data.frame(pca$rotation) %>%
  tibble::rownames_to_column('col') %>%
  dplyr::left_join(columns, by='col') %>%
  dplyr::mutate(col=factor(col, levels=columns %>% 
                             dplyr::filter(col %in% selected.cols) %>% 
                             dplyr::pull(col)),
                col_name=gsub("_downsampled|_switch",'',col))

@@

```{r Fig4J, fig.width=5.25 * fig_scale, fig.height=1.7 * fig_scale}
features.here <- c('frac_clusters_SG1',
                   'mean_cluster_size_downsampled',
                   'n_homology_switch',
                   'n_untemplated_switch',
                   'frac_breaks_intra',
                   'insert_frequency')

@@

ggplot(tmp, aes(x=ifelse(grepl('Cluster',group),as.numeric(group),1+as.numeric(group)),
                y=value,fill=group)) +
  geom_boxplot(outlier.shape=NA, lwd=.5) +
  geom_jitter(width=.1,height=0,size=.5) +
  #annotate('text',x=2,y=Inf,label='', size=2) +
  #annotate('text',x=6,y=Inf,label='validation', size=2) +
  # stat_compare_means(comparisons=list(c('Cluster1','Cluster3'),
  #                                     c('Cluster1','Cluster2'),
  #                                     c('Cluster2','Cluster3')),
  #                    method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  # stat_compare_means(comparisons=list(c('Control','ATM'),
  #                                     c('Control','BRCA1')),
  #                    method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(1,2),c(1,3),c(2,3)),
                     method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(5,6),c(5,7),c(6,7)),
                     method='wilcox.test',label = "p.signif", size=2, vjust=.5) +
  facet_wrap(.~col,ncol=length(features.here),scales='free_y') +
  my_theme() +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=setNames(c(rep('gray',3),rep('darkseagreen3',3)),groups)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') +
  scale_x_continuous(breaks=c(1,2,3,5,6,7),
                     labels=groups) +
  labs(x='',y='', color='Diagnosis')
```

@@

```{r FigS4B, fig.width=3 * fig_scale, fig.height=2.5 * fig_scale}
pvals <- list()
for (x in c("Tetanus",
            "Diphteria",
            "PnPS")) {
  for (f in selected.cols) {
    data <- FB.data.pooled[FB.samples,c(x,f)] %>%
      tidyr::drop_na()
    form <- as.formula(paste0(x,'=="negative" ~ ',f))
    res <- glm(form, family='binomial', data=data) %>% summary()
    pvals[[paste0(x,'_',f)]] <- data.frame(pval=coef(res)[2,4],
                                           vacc=x,
                                           col=f)
  }
}

@@

```{r inhibitors, fig.width=6.75 * fig_scale, fig.height=6.5 * fig_scale }
inhibitors_sample_sheet <- read.csv('../sodar/2019_Inhibitors/a_2019_Inhibitors.txt',
                                    sep='\t',header=1) %>%
  dplyr::left_join(read.csv('../sodar/2019_Inhibitors/s_2019_Inhibitors.txt',sep='\t',header=1),
                   by='Sample.Name') %>%
  dplyr::mutate(Donor=Source.Name,
                Treatment=Parameter.Value.Treatment.,
                Dose=Parameter.Value.Dose.,
                Batch=Parameter.Value.Batch.,
                Sample=Comment.Extract.Name.) %>%
  dplyr::distinct(Donor,Treatment,Dose,Sample,Batch,.keep_all = TRUE)

inhibitors.data <- read.csv('../human_samples/20240307_hg38_results/output/summary/20240307_human_samples_stats.csv',
                            header=1,row.names=1) %>%
  tibble::rownames_to_column('Sample') %>%
  dplyr::inner_join(inhibitors_sample_sheet %>% dplyr::distinct(Donor, Treatment, Sample, Batch), by='Sample') %>%
  dplyr::filter(nreads_final >= 500, frac_clusters_SA1+frac_clusters_SA2 > .05,
                frac_clusters_SG1 + frac_clusters_SG2 + frac_clusters_SG3 + frac_clusters_SG4 > .05) %>%
  dplyr::mutate_all(~replace_na(.,0)) %>%
  dplyr::mutate(Diagnosis=Treatment) %>%
  dplyr::select(tidyselect::any_of(c('Donor','Diagnosis','Sample','Batch',columns$col))) %>%
  dplyr::mutate(num_variants_somatic=num_variants * (1 - frac_variants_germline))

# inhibitors.X <- inhibitors.data[,selected.cols] %>% dplyr::mutate_if(is.numeric, scale)
# inhibitors.meta <- inhibitors.data[,c('Donor','Diagnosis')] %>%
#   dplyr::mutate(Diagnosis=ifelse(grepl('BIBR',Diagnosis),
#                                  'BIBR',
#                                  ifelse(grepl('ATM',Diagnosis),'ATMi',
#                                         ifelse(grepl('DMSO',Diagnosis),'DMSO',Diagnosis))))
# 
# pcadat <- cbind(prcomp(inhibitors.X)$x, inhibitors.meta)
# 
# ggplot(pcadat, aes(x=PC1,y=PC2)) +
#   geom_point(aes(color=Diagnosis), size=2) +
#   my_theme() +
#   scale_color_manual(values=c(diagnosis_colors,c('DMSO'='gray40','HU'='orange1','Olaparib'='blue2','SRC7'='orchid1',
#                                                  'DSMO'='gray40','SCR7'='orchid1','BIBR'='green3','ATMi'='brown4'))) +
#   guides(color=guide_legend(override.aes = list(shape=15,size=2), ncol=3)) +
#   theme(legend.key.height=unit(1.5,'mm'),
#         legend.key.spacing=unit(0,'mm'),
#         legend.position='bottom',
#         legend.box='horizontal') +
#   coord_cartesian(clip='off') +
#   labs(x='PC 1', #paste0('PC 1 (',100*signif(pca$sdev[1]**2/sum(pca$sdev**2),3),'%)'),
#        y='PC 2', #paste0('PC 2 (',100*signif(pca$sdev[2]**2/sum(pca$sdev**2),3),'%)'),
#        color='',shape='',fill='',alpha='')
# 
# pcadat.test <- rbind(cbind(as.matrix(human.X[train,]) %*% pca$rotation, human.meta[train,c('Donor','Diagnosis')]),
#                      cbind(as.matrix(human.X[test,]) %*% pca$rotation, human.meta[test,c('Donor','Diagnosis')]),
#                      cbind(as.matrix(inhibitors.X) %*% pca$rotation,inhibitors.meta)) %>%
#   dplyr::mutate(dataset=ifelse(grepl('FB',Donor),'FB','inhibitors'))
# 
# ggplot(pcadat.test, aes(x=PC1,y=PC2)) +
#   geom_polygon(data = pcadat.test %>% dplyr::filter(dataset=='FB', Diagnosis %in% c('CVID','Control')) %>%
#                  dplyr::group_by(Diagnosis) %>%
#                  dplyr::slice(chull(PC1, PC2)), alpha = 0.1, color=NA, aes(fill=Diagnosis), show.legend=FALSE) +
#   geom_point(aes(color=Diagnosis,alpha=dataset, shape=dataset), size=2) +
#   my_theme() +
#   scale_fill_manual(values=diagnosis_colors) +
#   scale_color_manual(values=c(diagnosis_colors,c('DMSO'='gray40','HU'='orange1','Olaparib'='blue2','SRC7'='orchid1',
#                                                  'DSMO'='gray40','SCR7'='orchid1','BIBR'='green3','ATMi'='brown4'))) +
#   scale_alpha_manual(values=c('FB'=.25,'inhibitors'=1)) +
#   scale_shape_manual(values=c('FB'=16,'inhibitors'=15)) +
#   guides(color=guide_legend(override.aes = list(shape=15,size=2), ncol=3),
#          shape=guide_legend(ncol=1),
#          alpha=guide_legend(ncol=1)) +
#   theme(legend.key.height=unit(1.5,'mm'),
#         legend.key.spacing=unit(0,'mm'),
#         aspect.ratio=pca$sdev[2]**2/pca$sdev[1]**2,
#         legend.position='bottom',
#         legend.box='horizontal') +
#   coord_cartesian(xlim=c(-8,10),
#                   ylim=c(-5,8),
#                   clip='off') +
#   labs(x='PC 1', #paste0('PC 1 (',100*signif(pca$sdev[1]**2/sum(pca$sdev**2),3),'%)'),
#        y='PC 2', #paste0('PC 2 (',100*signif(pca$sdev[2]**2/sum(pca$sdev**2),3),'%)'),
#        color='',shape='',fill='',alpha='')

features.here <- c('top_clone_occupancy_downsampled',
                   'mean_cluster_size_downsampled',
                   'cluster_gini_downsampled',
                   'mean_length',
                   'donor_complexity',
                   'donor_score_W',
                   'donor_score_WGCW',
                   'homology_fw',
                   'n_homology_switch',
                   'n_untemplated_switch',
                   'frac_breaks_single',
                   'frac_breaks_intra',
                   'frac_breaks_sequential',
                   'frac_breaks_inversions',
                   'frac_breaks_duplications',
                   'insert_frequency')

tmp <- cbind(rbind(CH12.data[,features.here],
                   FB.data.pooled[,features.here],
                   QP.data[,features.here],
                   inhibitors.data[,features.here]),
             rbind(CH12.meta %>%
                     dplyr::mutate(Diagnosis=Genotype,
                                   dataset='CH12') %>%
                     dplyr::select(Diagnosis, dataset),
                   human.meta[,c('Diagnosis','dataset')],
                   inhibitors.data[,c('Diagnosis'),drop=FALSE] %>%
                     dplyr::mutate(dataset='inhibitors',
                                   Diagnosis=ifelse(Diagnosis=='ATM','ATMi',Diagnosis)))) %>%
  dplyr::filter(Diagnosis %in% c('WT','Trp53bp1 KO','Brca1 KO','Rif1 KO','Lig4 KO','Control','CVID','TNFRSF13B','ATM','BRCA1',
                                 'DMSO','BIBR','ATMi')) %>%
  tibble::rownames_to_column('sample') %>%
  gather(col,value,features.here) %>%
  dplyr::mutate(col=factor(gsub("_downsampled|_switch",'',col),
                           levels=gsub("_downsampled|_switch","",features.here)),
                group=droplevels(factor(Diagnosis, levels=c(names(genotype_colors), names(diagnosis_colors),
                                                            c('DMSO','BIBR','ATMi')))))

ggplot(tmp, aes(x=ifelse(dataset=='CH12',as.numeric(group),
                         ifelse(dataset=='inhibitors',
                                as.numeric(group) + 2,
                                as.numeric(group) + 1)),
                y=value,fill=group)) +
  geom_boxplot(outlier.shape=NA, lwd=.5) +
  geom_jitter(width=.1,height=0,size=.1) +
  stat_compare_means(comparisons=list(c(1,2),
                                      c(1,3),
                                      c(1,4),
                                      c(1,5)),
                     method='wilcox.test',label = "p.signif", hide.ns=TRUE, size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(7,8),
                                      c(7,9),
                                      c(7,10),
                                      c(7,11)),
                     method='wilcox.test',label = "p.signif", hide.ns=TRUE, size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(13,14),
                                      c(13,15)),
                     method='wilcox.test',label = "p.signif", hide.ns=TRUE, size=2, vjust=.5) +
  facet_wrap(.~col,ncol=4,scales='free_y') +
  my_theme() +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=c(diagnosis_colors,genotype_colors,
                             c('DMSO'='gray40','HU'='orange1','Olaparib'='blue2','SCR7'='orchid1',
                               'BIBR'='green3','ATMi'='brown4','Act'='yellow3'))) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') +
  scale_x_continuous(breaks=c(seq(1,5), seq(1,5) + 6, seq(1,3) + 12),
                     labels=levels(tmp$group)) +
  labs(x='',y='', color='Diagnosis')
```
